--- 
title: "Final Project: Solar panel analysis"
author: "Group E"
output:
  html_document:
    fig_caption: yes
    toc: yes
    toc_float: yes
  pdf_document:
    fig_caption: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

<!-- Code chunk -->
```{r, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE);
library(data.table);
library(ggplot2);
library(outliers);
library(leaflet);
library(caret)
library(scales);
library(corrplot);

folder_path <- "/Users/vickylin/Desktop/MBD/R/workgroup_project"

dt <- readRDS(file.path(folder_path, "solar_dataset.RData"));
dt <- as.data.table(dt);

station_info <- read.table(file.path(folder_path, "station_info.csv"), sep = ",", header = TRUE);
station_info <- as.data.table(station_info);

```

<hr />

## EDA: preparation
<!-- Text and format chunk -->
First, we split the dataset in three sub-sets (**S1**, **S2** and **S3**) for further analysis.

- **S1**: Historic station production data
- **S2**: PCA data (upper part, up to Row *5113*)
- **S3**: PCA data (lower part, from Row *5114*)

```{r, echo = TRUE}
dt_s1_orig <- as.data.frame(na.omit(dt[, 2:99], invert=FALSE)); 
dt_s2_orig <- as.data.frame(dt[1:nrow(dt_s1_orig), 100:456]); 
dt_s3_orig <- as.data.frame(1+dt[nrow(dt_s2_orig):nrow(dt), 100:456]); 
```

<hr />

## [1.1] Compute Statistcs
<!-- Text and format chunk -->
We computed the same statistic measures for Sections S1, S2 and S3 separately.

The analysis of S1, being the most relevant output at this point, delivered the following output:

```{r, echo = FALSE}
station <- colnames(dt_s1_orig);
mean_station <- sapply(dt_s1_orig, mean, na.rm=TRUE);
sd_station <- sapply(dt_s1_orig, sd, na.rm=TRUE);
quantile_station <- sapply(dt_s1_orig, quantile, na.rm=TRUE);

statistics_s1 <- data.table(Station = c(station),
                 Mean = mean_station,
                 Sd = sd_station,
                 Min = quantile_station[1,],
                 Sec_qrt = quantile_station[2,],
                 Median = quantile_station[3,],
                 Fourth_qrt = quantile_station[4,],
                 Max = quantile_station[5,]
); 
```


```{r, echo = TRUE}
head(statistics_s1)
```

As mentioned before, we conducted the **same analysis** for S2 and S3 as well, resulting in datatables with the same respective structure.

<hr />

## [1.2] Outlier detection

To identify outliers in the dataset, we used a **1.5 * IQR** benchmark for Max and Min and we ran a *for*-loop.

```{r, include=FALSE, echo = FALSE}
dt_s1_orig.iqr <- dt_s1_orig

Outliers <- c()

for(i in 1:98){
  
  max <- quantile(dt_s1_orig.iqr[,i],0.75, na.rm=TRUE) + (IQR(dt_s1_orig.iqr[,i], na.rm=TRUE) * 1.5 )
  min <- quantile(dt_s1_orig.iqr[,i],0.25, na.rm=TRUE) - (IQR(dt_s1_orig.iqr[,i], na.rm=TRUE) * 1.5 )
  
  idx <<- which(dt_s1_orig.iqr[,i] < min | dt_s1_orig.iqr[,i] > max)
  
  print(paste(i, length(idx), sep=''))
  
  Outliers <<- c(Outliers, idx) 
}

Outliers
```

This identified that in the S1 dataset there are **0 outliers** within our defined benchmark. We therefore did not see the need to adjust the dataset.

<hr />

## [1.3] Statistical Visualization 

In this section, we first mapped the stations using *Leaflet*, next we plotted the *statistical analysis* output from above.

<hr />

##### Leaflet map

We clustered the stations by production mean:

- <span style="color:blue">**Blue**</span>: Mean >17m
- <span style="color:green">**Green**</span>: Mean between 16m and 17m
- <span style="color:red">**Red**</span>: Mean <16m

```{r, echo = FALSE}
map <- leaflet();
map <- addTiles(map);

for (i in 1:nrow(station_info)){
  map <- addCircleMarkers(map, lng = as.numeric(station_info[i,3]), lat = as.numeric(station_info[i,2]),
                          label=paste0(colnames(dt_s1_orig[i])," with production mean ",round(statistics_s1[i,2]/1000000, 1),"m"), 
                          labelOptions = labelOptions(noHide = F), radius = 7, 
                          color = if(as.numeric(round(statistics_s1[i,2]/1000000, 1)) < 16) {"red"}
                          else if (as.numeric(round(statistics_s1[i,2]/1000000, 1)) >= 16 & as.numeric(round(statistics_s1[i,2]/1000000, 1)) < 17) {"green"}
                          else {"blue"}
  )};
map;
```

<hr />

##### Group details {.tabset .tabset-fade .tabset-pills}

###### Blue

```{r, echo = FALSE}
blue_st <- c()
blue_mean <- c()

for (i in 1:nrow(station_info)){
  if(as.numeric(round(statistics_s1[i,2]/1000000, 1)) > 17){
    blue_st <- matrix(append(blue_st,statistics_s1[i,1]));
    blue_mean <- matrix(append(blue_mean,statistics_s1[i,2]))}
}

blue <- matrix(c(blue_st,blue_mean), ncol=2)
blue <- as.data.table(blue); setnames(blue, c("Station","Mean"))
```

```{r,}
head(blue)
nrow(blue)
```

Selected Example for further analysis: **GOOD**

<hr />

###### Green


```{r, echo = FALSE}
green_st <- c()
green_mean <- c()

for (i in 1:nrow(station_info)){
  if(as.numeric(round(statistics_s1[i,2]/1000000, 1)) >= 16 & as.numeric(round(statistics_s1[i,2]/1000000, 1)) < 17){
    green_st <- matrix(append(green_st,statistics_s1[i,1]));
    green_mean <- matrix(append(green_mean,statistics_s1[i,2]))}
}

green <- matrix(c(green_st,green_mean), ncol=2)
green <- as.data.table(green); setnames(green, c("Station","Mean"))
```

```{r,}
head(green)
nrow(green)
```

Selected Example for further analysis: **ACME**

<hr />


###### Red

```{r, echo = FALSE}
red_st <- c()
red_mean <- c()

for (i in 1:nrow(station_info)){
  if(as.numeric(round(statistics_s1[i,2]/1000000, 1)) < 16){
    red_st <- matrix(append(red_st,statistics_s1[i,1]));
    red_mean <- matrix(append(red_mean,statistics_s1[i,2]))}
}

red <- matrix(c(red_st,red_mean), ncol=2)
red <- as.data.table(red); setnames(red, c("Station","Mean"))
```

```{r,}
head(red)
nrow(red)
```

Selected Example for further analysis: **STIG**

<hr />


##### Visualization of S1

We first plotted all solar panels according to their daily production means. We highlighted high- and low-performers by labelling them accordingly with their station name.

```{r, echo = FALSE}
means_plot <- ggplot(statistics_s1, aes(x = statistics_s1$Station, y = statistics_s1$Mean));
means_plot <- means_plot + geom_point(color = 'red');
means_plot <- means_plot + ggtitle("Mean of Production by Station") + xlab("Station");
means_plot <- means_plot + theme(plot.title = element_text(face = "bold"));
means_plot <- means_plot + scale_y_continuous(name = "Mean (millions)", 
                                              breaks = waiver(), 
                                              labels = c("15","16","17","18","19"));
means_plot <- means_plot + theme(axis.text.x=element_text(angle = 90, vjust = 0.5, size = 5));
means_plot <- means_plot + geom_text(aes(label = ifelse(Mean > 18000000 | Mean < 15000000, as.character(Station),'')),
                                     hjust = 0, 
                                     vjust = 0, 
                                     size = 3);
means_plot <- means_plot + theme(panel.background = element_rect(fill = 'white'),
                                 panel.grid.minor = element_blank(),
                                 panel.grid.major = element_line(color = "black", linetype = "dotted"),
                                 panel.grid.major.x = element_blank(),
                                 axis.line = element_line(colour = "black"));
means_plot;
```

<hr />

Next, we also visualized the production standard deviations of each station. Again, we highlighted high- and low-performers by labelling them accordingly with their station name.

```{r, echo = FALSE}
sd_plot <- ggplot(statistics_s1, aes(x = statistics_s1$Station, y = statistics_s1$Sd));
sd_plot <- sd_plot + geom_point(color = 'red');
sd_plot <- sd_plot + ggtitle("Standard Deviation of Production by Station") + xlab("Station");
sd_plot <- sd_plot + theme(plot.title = element_text(face = "bold"));
sd_plot <- sd_plot + scale_y_continuous(name = "Standard Deviation (millions)", 
                                        breaks = waiver(), 
                                        labels = c("7.5","7.6","7.7","7.8","7.9","8.0","8.1"));
sd_plot <- sd_plot + theme(axis.text.x=element_text(angle = 90, vjust = 0.5, size = 5));
sd_plot <- sd_plot + geom_text(aes(label = ifelse(Sd > 8100000 | Sd < 7600000, as.character(Station),'')),
                               hjust = 0, 
                               vjust = 0, 
                               size = 3);
sd_plot <- sd_plot + theme(panel.background = element_rect(fill = 'white'),
                           panel.grid.minor = element_blank(),
                           panel.grid.major = element_line(color = "black", linetype = "dotted"),
                           panel.grid.major.x = element_blank(),
                           axis.line = element_line(colour = "black"));
sd_plot;
```

<hr />

```{r, echo = FALSE}
plot(density(10^(-7)*dt_s1_orig[,1]), 
     main = "Density distribution of Station's solar production", ylim =c(0,0.6),
     xlab = "Production per station in Joules (*10^7)");

for (i in 2:ncol(dt_s1_orig)){
  lines(density(10^(-7)*dt_s1_orig[, i]),col=i);
};
```

<hr />

##### Visualization of S2 & S3 {.tabset .tabset-fade .tabset-pills}

###### S2 Density

```{r, echo = FALSE}
plot(density(dt_s2_orig[,1]),
     main = "Density of Each PCI \n From 1994 to 2007",
     xlab = "Distribution of PCA values",xlim=c(-60,60),ylim=c(0,0.2));
for (i in 1:ncol(dt_s2_orig)){
  lines(density(dt_s2_orig[, i]),col=i);
};
```

<hr />

###### S3 Density

```{r, echo = FALSE}
plot(density(dt_s3_orig[,1]),
     main = "Density of Each PCI \n From 2008 to 2012-11-30",
     xlab = "Distribution of PCA values",xlim=c(-60,60),ylim=c(0,0.2));
for (i in 1:ncol(dt_s3_orig)){
  lines(density(dt_s3_orig[, i]), col = i)};
```

<hr />

## [1.4] Data Scaling

We furthermore scaled the data:

```{r, echo = TRUE}
tipify <- function(x){
  mu <- mean(x, na.rm = TRUE);
  s <- sd(x, na.rm = TRUE);
  s[s == 0] <- 1;
  x <- (x - mu) / s;
}
```

```{r, echo = FALSE}
df_s1_standard <- as.data.frame(sapply(dt_s1_orig,tipify));
df_s2_standard <- as.data.frame(sapply(dt_s2_orig,tipify));
df_s3_standard <- as.data.frame(sapply(dt_s3_orig,tipify));
```

And we plotted the resulting densities:

<hr />

##### S1

```{r, echo = FALSE}
plot(density(df_s1_standard[,1]),ylim=c(0,0.5), main = "Density of tipified solar panels' production",
     xlab = "Distribution of scaled values");
for (i in 2:ncol(df_s1_standard)){
  lines(density(df_s1_standard[, i]),col=i)
}
```

<hr />

##### Visualization of S2 & S3 {.tabset .tabset-fade .tabset-pills}

###### S2 Density

```{r, echo = FALSE}
plot(density(df_s2_standard[,1]),ylim=c(0,0.7),main = "Density of tipified predictors' values",
     xlab = "Distribution of scaled values");
for (i in 1:ncol(df_s2_standard)){
  lines(density(df_s2_standard[, i]),col=i)
}
```

<hr />

###### S3 Density

```{r, echo = FALSE}
plot(density(df_s3_standard[,1]),ylim=c(0,0.7),main = "Density of tipified predictors' values",
     xlab = "Distribution of scaled values");
for (i in 1:ncol(df_s3_standard)){
  lines(density(df_s3_standard[, i]),col=i)
}
```

<hr />


## [1.5] Correlation detection

We analyzed a) correlations between solar panels and respective PCAs and b) correlations between explanatory variables themsevles.

<hr />

##### Correlations between solar panels and PCA

We identified **PCA 1** to be the most relevant explanatory variable for 97 of the 98 solar panels.

The exception was station #60, MTHE, with PCA 2 as most relevant correlation.

```{r, echo = FALSE}
variable <- c();

for(iterator in 1:98){
  corr<-(as.data.table(cor(dt_s1_orig[,iterator],dt_s2_orig)));
  variable <- c(variable, max.col(corr))
}
corr_main <- as.data.table(statistics_s1[,1])
corr_main$Corr_PCA <- variable
```

```{r}
head(corr_main)
```


<hr />


##### Correlation between explanatory variables themselves

```{r, include=FALSE, echo = FALSE}
corr_PCA<-cor(dt_s2_orig);

#Remove the own correlation of variables
for (i in 1:length(colnames(corr_PCA))){
  corr_PCA[i,i]<-0
}
#obtain vector with the highest correlation per PCA to all other PCAs
max_corr_per_PCA<-c();
for (i in 1:length(colnames(corr_PCA))){
  a<-abs(max(abs(corr_PCA[,i])));
  max_corr_per_PCA<-append(max_corr_per_PCA,a)
};
```

```{r}
max(max_corr_per_PCA);
```

The highest absolute value of correlation coeficient of each PCA to other PCAs is 0,072.
Not enough evidence of correlation using this approach to make some PCAs redundant.


<hr />

## [1.6] Visualization of correlation 

To get a better understanding of above-conducted analysis we visualized the findings.

<hr />

##### Visualization solar panel to PCA correlation {.tabset .tabset-fade .tabset-pills}

###### Table

```{r, echo=FALSE}
input_corr<-cor(x=dt_s1_orig,y=df_s2_standard);
corrplot(input_corr[1:98,1:98],tl.cex = 0.3);
```

<hr />

###### Histogram

```{r, echo = FALSE}
variable_df <- as.data.frame(variable);
hist_fq <- ggplot(variable_df, aes(variable)) + geom_histogram(aes(y=(..count..)), binwidth = 1, color = 'white');
hist_fq <- hist_fq + ggtitle("Most Correlated PCA") + xlab("PCA") + ylab("Frequency (Sum = 98)"); #add titles
hist_fq <- hist_fq + theme(plot.title = element_text(face = "bold")); #change the face of main title
hist_fq <- hist_fq + theme(panel.grid.minor = element_blank(), 
                           panel.grid.major.x = element_blank(),
                           axis.line = element_line(colour = "black")); #remove unnecessary grid lines.
hist_fq <- hist_fq + scale_x_continuous(breaks = c(1,2),
                                        labels = c("PC1","PC2"));
hist_fq <- hist_fq + stat_bin(binwidth= 1, geom="label", aes(label=..count..));
hist_fq;
```



<hr />

##### Visualization PCA to PCA correlation {.tabset .tabset-fade .tabset-pills}

###### Table

```{r, echo=FALSE}
input_corr_PCA<-cor(dt_s2_orig);
corrplot(input_corr_PCA[1:50,1:50],order = "hclust",tl.cex = 0.5);
```

<hr />

###### Histogram

```{r, echo = FALSE}
max_corr_per_PCA_df <- as.data.frame(max_corr_per_PCA);
hist_cor <- ggplot(max_corr_per_PCA_df, aes(max_corr_per_PCA)) + geom_histogram(aes(y=(..count..)/sum(..count..)), binwidth = 0.005, color = 'white');
hist_cor <- hist_cor + ggtitle("PCA Correlation") + xlab("Maximum Correlation Coefficient Value") + ylab("Percentage of Total 357 PCA"); #add titles
hist_cor <- hist_cor + theme(plot.title = element_text(face = "bold")); #change the face of main title
hist_cor <- hist_cor + theme(panel.grid.minor = element_blank(), 
                             panel.grid.major.x = element_blank(),
                             axis.line = element_line(colour = "black")); #remove unnecessary grid lines.
hist_cor <- hist_cor + scale_y_continuous(labels = percent_format());
hist_cor;
```

<hr />

## [1.7] Dimensionality reduction

We used the following code to conduct the **Dimensionality reduction**:

```{r, echo=TRUE}
select_important <- function(dat, n_vars, y){
  varimp <- filterVarImp(x = dat, y=y, nonpara=TRUE);
  varimp <- data.table(variable=rownames(varimp),imp=varimp[, 1]);
  varimp <- varimp[order(-imp)];
  selected <- varimp$variable[1:n_vars];
  return(selected);
};
```

We conducted our analysis for three solar panels (Blue: *GOOD*, Green: *ACME*, Red: *STIG*) **representing the three groups identified in section [1.2]**. As the caclulation requires large amounts of CPU and memory (and therefore time), we post the output here instead.

- <span style="color:blue">**Blue [GOOD]**</span>
 <p><p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;[1] "PCA1"  "PCA2"  "PCA4"  "PCA5"  "PCA3"  "PCA6"  "PCA7"  "PCA24" "PCA9"  "PCA32"</p></p>

- <span style="color:green">**Green [ACME]**</span>: 
 <p><p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;[1] "PCA1"  "PCA2"  "PCA7"  "PCA4"  "PCA6"  "PCA5"  "PCA24" "PCA32" "PCA26" "PCA35"</p></p>

- <span style="color:red">**Red [STIG]**</span>:
 <p><p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;[1] "PCA1"  "PCA2"  "PCA7"  "PCA4"  "PCA5"  "PCA6"  "PCA24" "PCA26" "PCA32" "PCA33"</p></p>

<hr />

<span style="color:grey">Group E - Datadores</span>


```{js, echo=FALSE}
$('.title').css('color', 'Black')
$('.author').css('color', 'grey')
```


